---
title: "Translation of rmarkdown documents using a data frame"
author: "StatnMap"
date: "4 octobre 2017"
output:
  html_document: default
  pdf_document: default
lang: 'fr'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(knitr)
```

Simplfied version of blog article ["Translation of rmarkdown documents using a data frame"](//statnmap.com/2017-10-06-translation-rmarkdown-documents-using-data-frame) for direct use.

```{r lang}
# Define the output language
lang <- "fr"
```

```{r CreateTranslation}
# Create an external dataframe of translation
# can be replaced by a csv or excel file)
# Col-1 is original language ("en" here)
# Col-2 is translation
trans.df <- data.frame(
  en = c("Translated part", 
         "This is a long paragraph, with different ponctuations ! We can also add the content of a list in the document as follows:",
         "item 1",
         "item 2",
         "item 3", "subitem 3"),
  fr = c("Partie traduite", 
         "  Le contenu du    paragraphe  est le suivant : ",
         "élément 1",
         "élément 2", 
         "élément 3",
         "sous-élément 3"))

# Save it with a name that contains the lang attribute
readr::write_csv(trans.df, path = file.path(tempdir(), paste0(lang, ".csv")))

```

```{r ReadTranslation, message=FALSE}
# Read the external table file
trans.df <- readr::read_csv(file.path(tempdir(), paste0(lang, ".csv")))
# Remove space if before or after the text sequence
trans.df <- trans.df %>%
  mutate_all(function(.) gsub("[[:space:]]{2,}", " ", .)) %>%
  mutate_all(function(.) gsub("^[[:space:]]*|[[:space:]]*$", "", .)) %>%
  as.data.frame()
```

```{r TransRef}
# Translation reference
tr.orig <- trans.df[,"en"]
```

***

```{r MyFirstParagraph, results='asis', verbatim=TRUE}
cat('
# ', trans.df[tr.orig == "Translated part", lang], '

', trans.df[tr.orig == "This is a long paragraph, with different ponctuations ! We can also add the content of a list in the document as follows:", lang], '

- ', trans.df[tr.orig == "item 1", lang], '
- ', trans.df[tr.orig == "item 2", lang], '
- ', trans.df[tr.orig == "item 3", lang], '
    + ', trans.df[tr.orig == "subitem 3", lang], '
', sep = "")
```

***

```{r ListTranslationToDo}
# Read the present file
thisfile <- readr::read_lines("Template_translation_from_dataframe_simplified.Rmd")

# Find lines with translations
all_trans <- thisfile[grep('trans.df\\[tr.orig == "', thisfile)]
all_trans <- unique(all_trans)

# Extract translations
empty.file <- purrr::map(
  strsplit(all_trans, 'trans.df\\[tr.orig == "'), ~.x[2]) %>%
  purrr::map_chr(~strsplit(.x, '", lang]', fixed = TRUE)[[1]][1]) %>%
  data.frame(en = ., translation = "")

# Save as csv to send to translators
readr::write_csv(empty.file, path = "myemptyfile.csv")

```

After that, you can imagine everything you want:

- integrate more formatting like *italic*, **bold**, ...
- updating a table of translation to only send the missing translations to your translators
- use a table with multiple columns for all translations
- ...  

Be creative !  
You can also [contact me](//statnmap.com/contact/) for your projects like these or [any other project](//statnmap.com/freelance/) !

